### 2025-07-17 paper/algorithms/zetap.tex
More progress

\begin{algorithm}
\caption{Compute Position Error Indicator $\zeta_P$}
\label{alg:compute_zeta_p}
\begin{algorithmic}[1]
\REQUIRE Original contrast distribution $\boldsymbol{\chi}_o$
\REQUIRE Reconstructed contrast distribution $\boldsymbol{\chi}_r$
\ENSURE Position error indicator $\zeta_P$ (\%)

\STATE Extract magnitude information from both contrast distributions
\STATE Determine adaptive threshold based on reconstructed image dynamic range

\STATE Generate binary object masks using threshold-based classification
\STATE \quad - Original mask: pixels with non-zero contrast values
\STATE \quad - Reconstructed mask: pixels above computed threshold

\IF{no object regions detected in reconstruction}
    \RETURN complete failure indicator (100\%)
\ENDIF

\STATE Create normalized coordinate systems for both images
\STATE \quad - Map pixel coordinates to unit interval [0,1]
\STATE \quad - Handle potential resolution differences

\STATE Compute geometric centroids using weighted coordinate averaging:
\STATE \quad - Calculate center of mass for original object mask
\STATE \quad - Calculate center of mass for reconstructed object mask

\STATE Measure centroid displacement using Euclidean distance metric
\STATE Normalize distance to percentage scale relative to image dimensions

\RETURN position error as percentage of maximum possible displacement
\end{algorithmic}
\end{algorithm}