### 2025-07-17 paper/algorithms/zetas.tex

\begin{algorithm}
\caption{Compute Shape Error Indicator $\zeta_S$}
\label{alg:compute_zeta_s}
\begin{algorithmic}[1]
\REQUIRE Original contrast distribution $\boldsymbol{\chi}_o$
\REQUIRE Reconstructed contrast distribution $\boldsymbol{\chi}_r$
\REQUIRE Threshold factor $threshold = 0.5$ (default)
\ENSURE Shape error indicator $\zeta_S$ (\%)

\STATE Extract magnitude information from both contrast distributions
\STATE Determine adaptive threshold based on reconstructed image dynamic range
\STATE Apply binary classification to identify object and background regions

\IF{no object regions detected in reconstruction}
    \RETURN complete failure indicator (100\%)
\ENDIF

\STATE Extract object boundaries using contour detection methodology
\STATE Normalize spatial coordinates to handle resolution differences between images

\STATE Generate binary masks from contour information using polygon filling:
\STATE \quad - Apply odd-parity rule to determine interior points
\STATE \quad - Handle multiple contours for complex geometries

\STATE Compare geometric regions through symmetric difference operation
\STATE Quantify classification errors relative to original object size

\RETURN normalized shape error as percentage
\end{algorithmic}
\end{algorithm}